### 2020-01-19 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, ch10
statistical methods, ch10

How does this differ from the value estimated by MSBG?

#### 10.12
#### 10.13
#### 10.15
#### 10.16
#### 10.17
#### 10.18
#### 10.19
#### 10.20
#### 10.21
#### 10.22
#### 10.23
#### 10.24
#### 10.25
#### 10.26
#### 10.27
#### 10.28
#### 10.29
#### 10.30

>>>>

#### 10.12

For the following data, verify that ANOVA F test does not reject with $\alpha = 0.05$, but Welch's test _does reject_.

Group 1: 10, 11, 12, 9, 8, 7

Group 2: 10, 66, 15, 32, 22, 51

Group 3: 1, 12, 42, 31, 55, 19

```{r}
g1 <- c(10, 11, 12, 9, 8, 7)
g2 <- c(10, 66, 15, 32, 22, 51)
g3 <- c(1, 12, 42, 31, 55, 19)

data <- list(g1,g2,g3)

anova1(data) # p-value 0.082, fail to reject

t1way(data, tr=0) # p-value 0.046, reject
```

#### 10.13

Using the ANOVA F test, you see that $v_1 = 5$, and $v_2 = 8$. Verify that the number of groups is $J = 4$ and the total number of observations is $N = 12$.

$v_1$ = J - 1, J = 4

$v_2$ = N - J = 13

#### 10.14

A researacher reports a p-value of 0.001 with the ANOVA F test. Describe what conclusions are reasonable based upon this result.

_The distributions differ, suggesting that in particular the means differ._

#### 10.15

Summarize the reasons you might fail to reject with the ANOVA F test.

_Low power due to outliers, violating the equal variance assumption_

#### 10.16

Someone tests for equal variances and fails to reject. Does this justify the use of the ANOVA F test?

_No, this depends on power._

#### 10.17

A researcher reports that a test for normality was performed and that based on this test, no evidence of nonnormality was found. Why might it be unreasonable to assume normality despite this result?

_Unclear whether the test has enough power to detect a departure from normality._

#### 10.18

Outline how you might construct an example where sampling is from normal distributions, Welch's test rejects, but the ANOVA F test does not.

_Generate data so that groups have unequal variances._

#### 10.19

Consider a 2-by-2 design with population means:

$\mu_1 = 110, \mu_2 = 70$

$\mu_3 = 80, \mu_4 = 40$

```{r}
vals <- matrix(c(110, 70, 80, 40), nrow = 2, byrow = F)

apply(vals, 2, sum) # main effect

apply(vals, 1, sum) # interaction effect
```

#### 10.20

Consider a 2-by-2 design with population means.

$\mu_1 = 10, \mu_2 = 20$

$\mu_3 = 40, \mu_4 = 10$

```{r}
vals <- matrix(c(10, 20, 40, 10), nrow = 2, byrow = F)

apply(vals, 2, sum) # main effect

apply(vals, 1, sum) # interaction effect
```

#### 10.21

For the data in the last example of 10.1, use med1way to test hyhpothesis that population medians are equal.

```{r}
x=list()
x[[1]]=c( 7, 9, 8, 12, 8, 7, 4, 10, 9, 6)
x[[2]]=c(10, 13, 9, 11, 5, 9, 8, 10, 8, 7)
x[[3]]=c(12, 11, 15, 7, 14, 10, 12, 12, 13, 14)

med1way(x)
```

p < 0.05, reject

#### 10.22

Repeat previous with Qanvoa:

```{r}
Qanova(x)
```

p < 0.01, reject
